---
title: "Logistic Regression Modeling"
author: "Jose W. Martinez, Manuel A. Martinez, Jorge, Juan Martinez"
date: "1/31/2021"
output:
  html_document:
    github_document: true
    toc: true
    toc_float: true
    number_sections: false
    toc_depth: 5.0
    code_folding: hide
    code_download: true
    highlight: tango
    md_extensions: -autolink_bare_uris+hard_line_breaks
    includes:
    after_body: appendix.md
    linkcolor: cyan
bibliography: Epidemiology_COVID_19.bib
run_time: shiny
citecolor: blue
urlcolor: red
links-as-notes: true
biblio-title: References
link-citations: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = F, message = F, error = F)
library(rmdformats)
library(bookdown)
library(knitr)
library(dplyr)
library(tidyr)
library(e1071)
library(png)
library(tidyverse)
library(purrr)
library(cluster)
library(knitcitations)
library(knitLatex)
library(corrplot)
library(stringr)
library(dplyr)
library(ggplot2)
library(gganimate)
library(ggrepel)
library(GGally)
library(ggdendro)
library(reshape2)
library(cowplot)
library(grid)
library(anytime)
library(lubridate)
library(tibble)
library(kableExtra)
library(data.table)
library(htmlwidgets)
library(htmltools)
library(caret)
library(ISLR)
library(class)
library(plotly)
library(RColorBrewer)
library(ggsci)
library(gridExtra)
library(papeR)
library(mlbench)
library(epiDisplay)
library(Amelia)
library(readr)
library(hrbrthemes)
library(leaps)
library(scales)
library(gapminder)
library(tree)
library(lars)
library(dygraphs)
library(xts)          # To make the convertion data-frame / xts format
```

# Introduction

The novel coronavirus Sars-Cov2 (COVID-19) ... (Aqui siguen los médicos ...)

- Variable transformation:

```{r datainput, echo = T}

Covid_dt = read_csv("Covid20202021.csv")

attach(Covid_dt)

Covid_dt = Covid_dt[ , -c(3,4,5,6,7,9,28,30,33,66,68,69,70,71,76)]

Fec_not = lubridate::mdy(fec_not)
Covid_dt$fec_not = Fec_not

Fec_con = lubridate::mdy(fec_con)
Covid_dt$fec_con = Fec_con

Ini_sin = lubridate::mdy(ini_sin)
Covid_dt$ini_sin = Ini_sin

Fecha_Resultado = lubridate::mdy(FechaResultado)
Covid_dt$FechaResultado = Fecha_Resultado

Fec_hos = lubridate::mdy(fec_hos)
Covid_dt$fec_hos = Fec_hos

Fec_def = lubridate::mdy(fec_def)
Covid_dt$fec_def = Fec_def

Covid_dt$Defuncion = factor(Defuncion)
Covid_dt$cod_ase = factor(cod_ase)
Covid_dt$estrato = factor(estrato)
Covid_dt$Ciudad = factor(Ciudad)
Covid_dt$Sexo2 = factor(Sexo2)
Covid_dt$RangoEdad = factor(RangoEdad,
                            levels = c("0 a 9", "10 a 19", "20 a 29", "30 a 39",
                                       "40 a 49", "50 a 59", "60 a 69", "70 a 79",
                                       "80 a 89", "90 a 99"),
                            labels = c("0 a 9", "10 a 19", "20 a 29", "30 a 39",
                                       "40 a 49", "50 a 59", "60 a 69", "70 a 79",
                                       "80 a 89", "90 a 99"))
names(Covid_dt)[61] = c('Epidem_Week')



```

- Overview of the dataset:

```{r vartrans, echo = T, error = T, message=F, warning=F}


kbl(head(Covid_dt), booktabs = T,
    caption = 'Table 1. Overview of the dataset.') %>%
  kable_styling('hover', full_width = F) %>%
  scroll_box(height = '300px', width = '95%')
```

## Descriptive Analysis

### Total Number of Cases Over Time

```{r echo = T, error =F, message = F, warning=F}

don = Covid_dt %>%
  group_by(fec_not) %>%
  tally()

Don = xts(
  x = don$n, order.by = don$fec_not
)

dygraph(Don) %>%
  dyOptions(labelsUTC = TRUE, fillGraph=TRUE, fillAlpha=0.1, drawGrid = FALSE, colors="#D8AE5A") %>%
  dyRangeSelector() %>%
  dyCrosshair(direction = "vertical") %>%
  dyHighlight(highlightCircleSize = 5, highlightSeriesBackgroundAlpha = 0.2, hideOnMouseOut = FALSE)  %>%
  dyRoller(rollPeriod = 1)


```


### Number of Cases Over Time grouped by Sex

```{r echo = T, fig.height = 6, fig.width= 7, fig.cap='Figure 1. Total number of cases over time.'}

Cases_per_date = Covid_dt %>%
  mutate(Sex = ifelse((Sexo2 == 1) == T, 'Male', 'Female')) %>%
  group_by(fec_not, Sex) %>%
  tally()
  
fig1 = Cases_per_date %>%
  ggplot(aes(x = fec_not, y = n, group = Sex)) +
  geom_point(aes(colour = Sex), alpha = 0.5) +
  ylab('Total Number of Cases') + xlab('') +
  ggtitle('Cases per date') +
  theme(legend.position = 'bottom')


ggplotly(fig1, tooltip = c('n'))


```


```{r echo = T, fig.height = 6, fig.width= 7, fig.cap='Figure 2. Total number of cases grouped by Municipality.'}

Cases_by_Municipality = Covid_dt %>%
  mutate(Sex = ifelse((Sexo2 == 1) == T, 'Male', 'Female'),
         Municipality = factor(Ciudad)) %>%
  filter(Municipality %in% c('Pereira', 'Dosquebradas', 'Santa Rosa')) %>%
  group_by(Municipality, Sex, fec_not) %>%
  tally()
  
 
# Make a ggplot, but add frame=year: one image per year
ggplot(Cases_by_Municipality,
       aes(x = fec_not, y = n, group = Municipality)) +
  geom_point(aes(colour = Municipality), alpha = 0.5) +
  geom_line(aes(colour = Municipality), alpha = 0.5) +
  labs(title = 'Temporal Evolution of the Number of Positive Cases',
       x = '', y = 'Number of Cases') +
  transition_reveal(fec_not)

# Save at gif:
anim_save("271-ggplot2-animated-gif-chart-with-gganimate1.gif")
```

```{r echo = T, fig.height = 6, fig.width= 7, fig.cap='Figure 3. Total number of cases over time.'}


ggplot(Cases_per_date, aes(x = fec_not, y = n, group = Sex)) +
  geom_point(aes(colour = Sex), alpha = 0.5) +
  geom_line(aes(colour = Sex), alpha = 0.5) +
#  scale_color_viridis(discrete = TRUE) +
  labs(title = 'Temporal Evolution of the Number of Positive Cases',
        x = '', y = 'Number of Cases') +
#  theme_ipsum() +
  ylab("Number of babies born") +
  transition_reveal(fec_not)



# Save at gif:
anim_save("287-smooth-animation-with-tweenr.gif")
```



### Number of Cases Over Time grouped by Municipality

```{r echo = T, fig.height = 15, fig.width= 7, fig.cap='Figure 4. Number of cases per date grouped by municipality.'}

Covid_dt %>%
  mutate(Sex = ifelse((Sexo2 == 1) == T, 'Male', 'Female'),
         Municipality = Ciudad) %>%
  group_by(Municipality, fec_not, Sex) %>%
  tally() %>%
  ggplot(aes(x = fec_not, y = n, group = Sex)) +
  geom_point(aes(colour = Sex), alpha = 0.25) +
  ylab('Number of Cases') + xlab('') +
  facet_wrap(~Municipality, ncol = 2) +
  ggtitle('Cases per date') +
  theme(legend.position = 'bottom')


```

### Total Number of Cases grouped by Age Group


```{r echo = T, fig.height = 9, fig.width= 6, fig.cap='Figure 5. Number of cases per date grouped by municipality.'}

Covid_dt %>%
  filter(Ciudad %in% c('Pereira', 'Dosquebradas', 'Santa Rosa', 'La Virginia')) %>%
  mutate(Group = RangoEdad,
         Municipality = Ciudad) %>%
  group_by(Municipality, fec_not, Group) %>%
  tally() %>%
  ggplot(aes(x = fec_not, y = n, group = Group)) +
  geom_line(aes(colour = Group), alpha = 0.25) +
  ylab('Number of Cases') + xlab('') +
  facet_wrap(~Municipality, nrow = 4) +
  ggtitle('Cases per date') +
  theme(legend.position = 'bottom')


```

### Total Number of Cases grouped by Age Group and Sex


```{r fig.height= 5, fig.width=5, message = F, warning = F, error = F, fig.cap = 'Figure 6. 3D representation of  the number of positive cases over time.'}

Cases_per_Group_Sex = Covid_dt %>%
  mutate(Group = RangoEdad,
         Sex = ifelse((Sexo2 == 1) == T, 'Male', 'Female')) %>%
  group_by(Group, Sex, fec_not) %>%
  tally()

axx <- list(
  title = "Sex"
)

axy <- list(
  title = "Date"
)

axz <- list(
  title = "Number of Cases"
)

Num_Cases = Cases_per_Group_Sex$n
Sex = Cases_per_Group_Sex$Sex
Date = Cases_per_Group_Sex$fec_not

plot_ly(z = ~Num_Cases, x = ~Sex, y = ~Date,
        size = I(4)) %>%
  layout(scene = list(xaxis=axx,yaxis=axy,zaxis=axz)) 

```


### Total Number of Cases by EPS


```{r echo = T, fig.height = 11,fig.width= 5, fig.cap='Figure 7. Number of cases per healthcare provider (EPS).'}

Covid_dt %>%
  mutate(Sex = ifelse((Sexo2 == 1) == T, 'Male', 'Female'),
         Municipality = Ciudad) %>%
  group_by(Municipality, cod_ase, Sex) %>%
  tally() %>%
  ggplot(aes(y = cod_ase, group = Sex)) +
  geom_bar(aes(colour = Sex, fill = Sex), alpha = 0.25) +
  ylab('EPS') + xlab('Number of Cases') +
  ggtitle('Number of Cases by EPS') +
  theme(legend.position = 'bottom')


```



```{r fig.height= 17, fig.width=7}

fig_eps = Covid_dt %>%
  filter(Ciudad == 'Pereira') %>%
  mutate(Municipality = Ciudad,
         EPS = factor(cod_ase)) %>%
  group_by(Municipality, fec_not, EPS) %>%
  tally() %>%
  ggplot(aes(x = fec_not, y = n, group = EPS)) +
  geom_line(aes(colour = EPS), alpha = 0.35, linetype = 'dashed') +
  ylab('Number of Cases') + xlab('') +
  ggtitle('Number of Cases over time grouped by EPS') +
  theme(legend.position = 'none')

ggplotly(fig_eps, tooltip = c('n', 'EPS'))
```



# Methodology

We used 80% of the dataset for training, and 20% for testing. Also, we applied a 10-fold cross-validation to train and select the models. Since there were an important number of `NA`s, we removed the following variables:

* `estrato`
* `con_fin`
* `trab_salud`
* `deter_clin`
* `fec_hos`
* `fec_def`
* `Epidem_Week`
* `otros_dc`
* `TPRUEBA`
* `Edad`
 

The following chunk describes how the dataset was subdivided into two `dataframes`using a random subdividision of the data considering the response variable (`Defuncion`) as reference:

* `Covid_dt_train`
* `Covid_dt_train`

```{r echo =T}

#Code_Ase = ifelse(is.na(cod_ase) == T, 888888, cod_ase)

#Covid_dt$cod_ase = Code_Ase

#set.seed(1234)
#train <- createDataPartition(Covid_dt$Defuncion, p = 0.8, list = F, times = 1)

Covid_dt_train = Covid_dt[train, -c(1,2,5,18,19,20,21,22,23,24,25,49,55,56,58,59,60,61,62,63,65,67,68)]


write.csv(Covid_dt_train, 'Covid_dt_train.csv', sep = ',', row.names = F)
Covid_dt_test = Covid_dt[-train,-c(1,2,5,18,19,20,21,22,23,24,25,49,55,56,58,59,60,61,62,63,65,67,68)]

write.csv(Covid_dt_test, 'Covid_dt_test.csv', sep = ',', row.names = F)

head(Covid_dt_train)

```

## Validation Set Approach

```{r}

logistic.train.model <- glm(Defuncion ~ .,
                            family = binomial,
                            data = df)

model_train_set = summary(logistic.train.model)
Coef_assessment_train_set = prettify(summary(logistic.train.model))
        
train_set_probs = predict(logistic.train.model, Covid_dt_test, type = "response")
train_set_class = ifelse(train_set_probs > 0.5, 1, 0)
test_error_train_set = round(100*mean(train_set_class != Covid_dt_test$Defuncion), 2)

conf_mat_train_set = table(train_set_class[[i]], Covid_dt_test$Defuncion)
```



## Parameter Tuning

The `fitControl()` function performs a 10-fold cross-validation:

```{r echo = T}
#------------------------------------------------------------------------------
# GLM Cross Validation: 

K = 10

set.seed(1234)
rand_ind <- groupKFold(c(1:dim(Covid_dt_train)[1]), k = K)

sm <- list()
model_summary <- list()
model <- list()
conf_mat <- list()
Coef_assessment <- list()

k_fold_probs <- list()
k_fold_class <- list()
test_error = 0
  
for(i in 1:K){
  
  df = Covid_dt_train[rand_ind[[i]],]
  df_test = Covid_dt_train[-rand_ind[[i]],]
  
  logistic.fit = glm(Defuncion ~ .,
                      family = binomial,
                      data = df)
        
  model[[i]] = logistic.fit
  model_summary[[i]] = summary(logistic.fit)
  Coef_assessment[[i]] = prettify(summary(logistic.fit))
        
  k_fold_probs[[i]] = predict(logistic.fit, df_test, type = "response")
  test_error[i] = round(100*mean(k_fold_probs != df_test$Defuncion), 2)
        
  #k_fold_class[[i]] = ifelse(k_fold_probs > 0.5, 1, 0)
        
  conf_mat[[i]] = table(k_fold_class[[i]], df_test$Defuncion)
}
      
      CV_summary = data.frame(cbind(
        'Fold' = (1:kc),
        'Test Error' = test_error))
      
#      low_error = which(min(test_error) == T, arr.ind = T)
      
#      Lowest_test_error = test_error[low_error]
      
#      Model = model[[low_error]]
#      Model_Summary = model_summary[[low_error]]
#      k_fold_Classification = k_fold_class[[low_error]]
#      Confusion_Matrix = conf_mat[[low_error]]
      
      performance <- list("Model" = model,
                          "k-fold CV summary"  = CV_summary,
                          "Model Summary" = model_summary,
                          "Coefficients" =  Coef_assessment,
                          "Predicted outcome" = k_fold_probs,
                          "Confusion Matrix" = conf_mat)
      
      return(performance)
}



Logistic_Model = k_fold_cv_logistic(K, Covid_dt_train, Covid_dt_test)

```


## Logistic Regression Model

Logistic regression are models are often considered for this type of situations (*we neeed to add references here!*), and thus we consider this model as a reference to assess the accuracy of the other models.

### Full Model

(*We still can add more regressors*)

#### Training

The function `train()` trains the model using the pre-specified data partition:

```{r echo = T}
# 'LogitBoost'

set.seed(12345)
full_model_logistic <- train(Defuncion ~ semana + edad + gp_discapa + gp_gestan +
                             Riesgos + tos + fiebre + odinofagia + dif_res + adinamia +
                             rinorrea + conjuntivi + cefalea + diarrea + perd_gusto +
                             asma + epoc + Comorbilidades + Sexo2,
                             data = Covid_dt_train,
                             method = "Logit", 
                             trControl = fitControl)

full_model_logistic


ggplot(full_model_logistic)

whichTwoPct <- tolerance(full_model_logistic$results, metric = "Accuracy", 
                         tol = 2, maximize = TRUE)  
#cat("best model within 2 pct of best:\n")

full_model_logistic$results[whichTwoPct,]

full_model_logistic_pred = predict(full_model_logistic, newdata = Covid_dt_test)

Conf_Mat_full_model_logistic = table(full_model_logistic_pred, Covid_dt_test$Defuncion)
Conf_Mat_full_model_logistic
#sensit_full_model_Logistic <- Conf_Mat_full_model_logistic[2,2]/(Conf_Mat_full_model_logistic[1, 2] + Conf_Mat_full_model_logistic[2, 2])

#sensit_full_model_Logistic = round(100*sensit_full_model_Logistic, 2)

#ROC_test_full_model_logistic <- data.frame(matrix(vector(),
#                                            nrow = dim(Covid_dt_test)[1],
#                                            ncol = 4,
#                                            dimnames = list(c(),
#                                                            c("obs",
#                                                              "pred",
#                                                              "Yes",
#                                                              "No"))),
#                                           stringsAsFactors = T)

#levels(Covid_dt_test$Defuncion) <- c("Survived", "Death")
#Failure_Pred_MMR_Liberal <- factor(Failure_Pred_MMR_Liberal)
#levels(full_model_logistic_pred) <- c("Survived", "Death")

#ROC_test_full_model_logistic$obs <- Covid_dt_test$Defuncion
#ROC_test_full_model_logistic$pred <- full_model_logistic_pred
#ROC_test_full_model_logistic$Yes <- rep(length(Covid_dt_test$Defuncion[Covid_dt_test$Defuncion == "Death"])/length(Covid_dt_test$Defuncion))
#ROC_test_full_model_logistic$No <- 1 - ROC_test_full_model_logistic$Yes

#twoClassSummary(ROC_test_full_model_logistic,
#                lev = levels(ROC_test_full_model_logistic$obs))
```
According to this output, the model has a high accuracy to predict the `Survival` of the patients, but a low accuracy to predict `Death` of a patient.

## Implicit Feature Selection

```{r echo = T, eval=FALSE}
set.seed(12345)
full_model_AdaBoost <- train(Defuncion ~ semana + edad + gp_discapa + gp_gestan +
                             Riesgos + tos + fiebre + odinofagia + dif_res + adinamia +
                             rinorrea + conjuntivi + cefalea + diarrea + perd_gusto +
                             asma + epoc + Comorbilidades + Sexo2,
                             data = Covid_dt_train,
                             method = "adaboost", 
                             trControl = fitControl)
```




## Random Forest


## ROC Curves

## Neural Networks

### Multi-Layer Perceptron

```{r echo = T, eval = T}
set.seed(12345)
full_model_MLP <- train(Defuncion ~ semana + edad + gp_discapa + gp_gestan +
                          Riesgos + tos + fiebre + odinofagia + dif_res + adinamia +
                          rinorrea + conjuntivi + cefalea + diarrea + perd_gusto +
                          asma + epoc + Comorbilidades + Sexo2,
                        data = Covid_dt_train,
                        method = "mlp", 
                        trControl = fitControl)

full_model_MLP_pred = predict(full_model_MLP, newdata = Covid_dt_test)

Conf_Mat_full_model_MLP = table(full_model_MLP_pred, Covid_dt_test$Defuncion)
Conf_Mat_full_model_MLP
```


# References {-}

<div id="refs"></div>
